---
title: "Summarizing Relationships, Tidying and Joining Data"
date: 2023-03-29
---


## Readings 

- For summarizing relationships, see QSS 3.6 and MD 5.1.1. 
- For pivoting and tidy data, see MD Ch 4. 
- For more on joins, see the [chapter 20](https://r4ds.hadley.nz/joins.html) in R for Data Science.

## Data

- Reinstall the TPDstat data package if the current package does not have the dataset(s) from the lecture 

```{r}
#| eval: false
remotes::install_github("seungho-an/TPDdata")
```

<!-- - If you cannot install the package, you can download the data locally at [`covid_votes.csv`](/files/data/covid_votes.csv)  to the same directory as your Rmd file and load it using: -->
<!-- ```{r} -->
<!-- #| eval: false -->
<!-- library(tidyverse) -->
<!-- covid_votes <- read_csv("covid_votes.csv") -->
<!-- ``` -->

<!-- - For Thursday lecture, we added to additional data sets to `gov50data`. If you cannot install that package, then you can download them: [`mortality.csv`](/files/data/mortality.csv) and [`spotify.csv`](/files/data/spotify.csv). Put them in the same folder as your Rmd file and use these commands to load: -->
<!-- ```{r} -->
<!-- #| eval: false -->
<!-- library(tidyverse) -->
<!-- mortality <- read_csv("mortality.csv") -->
<!-- spotify <- read_csv("spotify.csv") -->
<!-- ``` -->

- We will also use data from the `nycflights13` package that you can install with:
```{r}
#| eval: false
install.packages("nycflights13")
```




## Slides and Code

- Summarizing relationships, writing our own functions, causality review, pivoting longer, joining data

  - [PDF slides](/slides/05_bivariate_tidying.pdf)
  - [PDF handout version](/slides/05_bivariate_tidying_handout.pdf) (no incremental slides)
  - Code: [`covid_votes.Rmd`](/files/code/covid_votes.Rmd)
  - Code: [`lecture-05.Rmd`](/files/code/lecture-05.Rmd)


